---
title: "Allometry in LUMCON marshes"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include = FALSE, echo = FALSE}
### This section loads libraries, sets some local variables

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>" 
)

if(!require(knitr)){
  install.packages("knitr", repos='http://cran.us.r-project.org')
}
if(!require(rmarkdown)){
  install.packages("rmarkdown", repos='http://cran.us.r-project.org')
}
if(!require(plyr)){
  install.packages("plyr", repos='http://cran.us.r-project.org')
}
if(!require(reshape2)){
  install.packages("reshape2", repos='http://cran.us.r-project.org')
}
if(!require(ggplot2)){
  install.packages("ggplot2", repos='http://cran.us.r-project.org')
}
if(!require(gridExtra)){
  install.packages("gridExtra", repos='http://cran.us.r-project.org')
}
if(!require(scales)){
  install.packages("scales", repos='http://cran.us.r-project.org')
}
if(!require(MuMIn)){
  install.packages("MuMIn", repos='http://cran.us.r-project.org')
}
if(!require(rsq)){
  install.packages("rsq", repos='http://cran.us.r-project.org')
}

if(!require(multcomp)){
  install.packages("multcomp", repos='http://cran.us.r-project.org')
}
if(!require(sandwich)){
  install.packages("sandwich", repos='http://cran.us.r-project.org')
}
if(!require(car)){
  install.packages("car", repos='http://cran.us.r-project.org')
}
if(!require(Rmisc)){
  install.packages("Rmisc", repos='http://cran.us.r-project.org')
}
if(!require(tidyverse)){
  install.packages("tidyverse", repos='http://cran.us.r-project.org')
}
if(!require(lattice)){
  install.packages("lattice", repos='http://cran.us.r-project.org')
}
if(!require(nlme)){
  install.packages("nlme", repos='http://cran.us.r-project.org')
}
if(!require(lubridate)){
  install.packages("lubridate", repos='http://cran.us.r-project.org')
}
# if(!require(svglite)){
#   install.packages("svglite", repos='http://cran.us.r-project.org')
# }

library(ggplot2)
library(gridExtra)
library(plyr)
library(LUMCONplants)
library(multcomp) # for glht
library(sandwich) # for HC3 estimator
library(car) # for Levene's test

library(Rmisc)
library(tidyverse)
library(lattice)
library(nlme)
library(lubridate)
# library(svglite) # doesn't install properly on linux


### define some constants
plotSize       <- 0.25 * 0.25  # size of quadrat, m2
coreTube       <- pi*(6.9/2)^2 # area of plastic coring tube used prior to August 2016: units = cm2 
coreTube_metal <- pi*(6.5/2)^2 # area of metal coring device used beginning in August 2016: units = cm2 
growingMonths  <- 4:10         # growing season defined as April:October
bagMass        <- 5.4          # grams; bags used for belowground cores
tckSize        <- 0.01

getDemographics <- function(data = stemDat, status = "LIVE", groupBreaks = c(0, 25, 50, 100, 200), groupLabels = c("0-25 cm", "25-50 cm", "50-100 cm", ">100 cm"), quadrat_m2 = 1) {
  df1 <- transform(data[(data$status %in% status), ], group = cut(hgt, 
             breaks = groupBreaks,
             labels = groupLabels))
  res <- do.call(data.frame, aggregate(hgt ~ time + group, df1, 
             FUN = function(x) c(count = sum(!is.na(x)) / quadrat_m2, mean = mean(x, na.rm = TRUE) / quadrat_m2)))
  dNew <- data.frame(group = levels(df1$group))
  df2 <- merge(res, dNew, all = TRUE)
  
  # add missing bins, set to 0 hgt/dens
  missedLevs <- expand.grid(time = unique(res$time), group = unique(res$group), hgt.count = 0, hgt.mean = 0)
  missingLevs <- missedLevs[which(!paste0(missedLevs$time, missedLevs$group) %in% paste0(res$time, res$group)), ]
  df2 <- rbind(df2, missingLevs)
  #df2$site <- targReg
  invisible(df2)
}


regionLabel <- function(data, siteCol = "site") {
  # function takes a dataset and the name of the column with site names (e.g., "LUM1"), and 
  # adds a column with marsh names ("LUM", "TB-A", "TB-B")
  data$region <- as.character(NA)
  for (i in 1:nrow(data)) {
    if (data[i, siteCol] %in% paste0("LUM", 1:3)) {
      data$region[i] <- "LUMCON"
    } else if (data[i, siteCol] %in% c("TB1", "TB2")) {
      data$region[i] <- "Bay LaFleur"
    } else if (data[i, siteCol] %in% c("TB3", "TB4")) {
      data$region[i] <- "Lake Barre"
    }
  }
  data
}


### a function to make correlation matrices with significance stars
### from http://myowelt.blogspot.com/2008/04/beautiful-correlation-tables-in-r.html
### usage: kable(corstarsl(as.matrix(data))) 
corstarsl <- function(x){
  x <- as.matrix(x)
  R <- Hmisc::rcorr(x)$r
  p <- Hmisc::rcorr(x)$P
  
  ## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
  
  ## trunctuate the matrix that holds the correlations to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
  
  ## build a new matrix that includes the correlations with their apropriate stars
  Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
  diag(Rnew) <- paste(diag(R), " ", sep="")
  rownames(Rnew) <- colnames(x)
  colnames(Rnew) <- paste(colnames(x), "", sep="")
  
  ## remove upper triangle
  Rnew <- as.matrix(Rnew)
  Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
  Rnew <- as.data.frame(Rnew)
  
  ## remove last column and return the matrix (which is now a data frame)
  Rnew <- cbind(Rnew[1:length(Rnew)-1])
  return(Rnew)
} 


```


```{r allometry models, echo = FALSE, include = FALSE, message = FALSE}
### This section parameterizes allometry models and uses the models to estimate unknown stem masses

stemDat$moYr  <- paste0(stemDat$mo, "-", substr(stemDat$yr, 3, 4))
stemDat$time  <- as.yearmon(stemDat$moYr, "%b-%y")
stemDat       <- seasonLabel(data = stemDat)
stemDat$seas  <- substr(stemDat$season, 1, 4)
stemDat$plantYr <- as.numeric(paste0(20, substr(stemDat$season, 6, 7))) # avoids splitting winter samples (Jan 2018 is in winter 2017)


### separate plot As for allometry work
stm <- stemDat[stemDat$plot %in% "A", ]



### allometry models
# nls(mass_tot ~ I(a * hgt^b), start = list(a = 0.2, b = 0.2), data = stm[(stm$site %in% "LUM1") & (stm$status %in% "LIVE") & (stm$seas %in% "fall"), ])
# models <- dlply(stm, .(site, season, status), function(x) 
#   nls(mass_tot ~ I(a * hgt^b), start = list(a = 0.2, b = 0.2), data = x)) 
# ldply(models, coef)

models.hiRes <- ddply(stm, .(site, seas), function(x) 
  getParams(dataset = x))
models.reg <- ddply(stm, .(region, seas, plantYr), function(x) 
  getParams(dataset = x))

models            <- ddply(stm, .(region, seas), function(x) 
  getParams(dataset = x))
dead.allom        <- getParams(dataset = stm)
dead.allom$region <- "all"
dead.allom$seas   <- "all"

models <- rbind.fill(models, dead.allom)


exponents_vs_coefs <- ggplot(aes(y = exp.live, x = coef.live, colour = seas), data = models) + geom_point() + theme_classic() + labs(y = "exponent", x = "coefficient")
# spring = low coefficients, high exponents


allom_by_seas_plantYr <- ggplot(aes(y = mass_tot, x = hgt, colour = site), 
       data = stm[(stm$region %in% "LUMCON") & (stm$status %in% "LIVE"), ]) + 
  geom_point() + theme_classic() + labs(y = "mass", x = "hgt") + 
  facet_grid(seas ~ plantYr)

allom_by_seas <- ggplot(aes(y = mass_tot, x = hgt, colour = site), 
       data = stm[(stm$region %in% "LUMCON") & (stm$status %in% "LIVE"), ]) + 
  geom_point() + theme_classic() + labs(y = "mass", x = "hgt") + 
  facet_grid(seas ~ .)




# Apply allometry to un-weighed stems -------------------------------------

### Generate estimates from season-region models and pooled data 
for (i in 1:nrow(stemDat)) {
  # identify region and season
  targRegion <- stemDat$region[i]
  targetSeas <- stemDat$seas[i]
  
  # find seasonal allometry parameters for the region
  if (stemDat$status[i] %in% "LIVE") {
    coefficient <- models$coef.live[(models$region %in% targRegion) & (models$seas %in% targetSeas)]
    exponent    <- models$exp.live[(models$region %in% targRegion) & (models$seas %in% targetSeas)]
  } else if (stemDat$status[i] %in% "DEAD") {
    coefficient <- models$coef.dead[(models$region %in% "all") & (models$seas %in% "all")]
    exponent    <- models$exp.dead[(models$region %in% "all") & (models$seas %in% "all")]
  }
  # apply allometry to get mass estimate
  stemDat$mass.est1[i] <- coefficient * (stemDat$hgt[i] ^ exponent) # season-region models for live, pooled model for dead
  if (stemDat$status[i] %in% "LIVE") {
    stemDat$mass.est2[i] <- models$coef.live[(models$region %in% "all")] * (stemDat$hgt[i] ^ models$exp.live[(models$region %in% "all")]) # single model for all data (always used for dead stems)
  } else {
    stemDat$mass.est2[i] <- NA
  }
  
  # define final mass to be used
  # set dummy variable if plant mass was actually measured
  if (!is.na(stemDat$mass_tot[i])) {
    stemDat$modeled[i] <- 0
    stemDat$mass[i]    <- stemDat$mass_tot[i]
  } else {
    stemDat$modeled[i] <- 1
    stemDat$mass[i]    <- stemDat$mass.est1[i]
  }
}


```




```{r allometry ANOVAs, echo = FALSE, include = FALSE, message = FALSE}
### Identify differences between seasons in the LUMCON region (pooled)
subData       <- stm[(stm$status %in% "LIVE") & !is.na(stm$hgt) & !is.na(stm$mass_tot) & (stm$region %in% "LUMCON"), ]
subData$seas  <- ordered(subData$seas, c("sprg", "sumr", "fall", "wint"))
full.nls <- nls(mass_tot ~ I(a * hgt^b), data = subData, start = list(a = 0.2, b = 0.2))
summary(full.nls)

subData$resid <- resid(full.nls)

# Residuals have unequal variances among seasons (based on Levene Test and boxplots). 
# To make comparisons I follow the procedure described in Herberich et al. 2010
leveneTest(resid ~ seas, data = subData) 
boxplot(resid ~ seas, data = subData)

# ANOVA of residuals by seas
modSeas          <- aov(resid ~ seas, data = subData)
summary(modSeas) # F = 104.9 p<<0.001   DEAD: F = 3.2; p = 0.02

# glht sets up multiple contrasts. vcov = vcovHC specifies heteroscedasticity-consistent estimator of covariance matrix (Herberich et al. 2010)
modSeas_glht     <- glht(modSeas, linfct = mcp(seas = "Tukey"), vcov = vcovHC)
coef(modSeas_glht)
summary(modSeas_glht) 
plot(confint(modSeas_glht)) # all seasons are significantly different. Dead stem allometry: no seasonal differences


```


```{r trends in stem demographics, echo = FALSE, include = FALSE, message = FALSE}
### This section explores trends in stem size class distributions
### get demographics by plot
int_df  <- ddply(stemDat, .(region, site, plot), function(x) 
  getDemographics(x, quadrat_m2 = plotSize))

### mean by marsh
final_df <- ddply(int_df, .(region, site, group, time), summarise, 
                  stmDens    = mean(hgt.count, na.rm = TRUE),
                  stmDens.se = se(hgt.count),
                  hgt        = mean(hgt.mean, na.rm = TRUE),
                  hgt.se    = se(hgt.mean))

stm.demographics <- ggplot(final_df[final_df$region %in% "LUMCON",], 
       aes(y = stmDens, x = as.numeric(time))) + #, col = region)) + 
  geom_point(size = 1.2) + geom_errorbar(aes(ymin = (stmDens - stmDens.se), ymax = (stmDens + stmDens.se)), width = 0) +
  geom_line(lwd = 0.7) +
  theme_classic() + theme(legend.title = element_blank()) + 
  ylab(expression("Live stem density (m"^-2~")")) + xlab("") + 
  facet_wrap("group", nrow = 2) + 
  theme_bw() %+replace% theme(strip.background  = element_blank(), legend.title = element_blank())


### add in plot-level mass data
mass_df <- ddply(stemDat, .(region, site, plot, time), summarise, 
                 mass = mean(mass, na.rm = TRUE),
                 mass.se = se(mass))
mass_df2 <- join_all(list(mass_df, int_df))

ggplot(mass_df2, aes(x = log(hgt.count), y = log(mass))) + geom_point() + theme_classic()



# identify seasonality
for (i in 1:length(unique(final_df$group)[1:3])) {
  test <- ts(final_df[(final_df$region %in% "LUMCON") & 
                        (final_df$group %in% unique(final_df$group)[i]), "stmDens"], 
            start = c(12, 5), deltat = 1/12)
  test2 <- stl(test, s.window = 12)
  print(apply(test2$time.series, 2, var) / var(test))
  
  # evaluate change as the difference between start and end values of the trend component 
  print(noquote(paste0("trend for stems in ", unique(final_df$group)[i], " size class: ",
                       round(as.numeric(test2$time.series[nrow(test2$time.series), 2] - test2$time.series[1, 2])))))
}
 # no obvious trends

```




## Allometry data

```{r allometry by season-year, fig.width = 6, fig.height = 6, echo=FALSE, warning=FALSE, fig.cap = "Mass-height allometry at LUMCON marshes, by season (rows) and year (columns)."}

allom_by_seas

```

Allometry models differ significantly by season, and seasonality has become more pronounced since Hill and Roberts (2017). 


```{r allometry plot 2, fig.width = 3.5, fig.height = 2.5, warning=FALSE, echo=FALSE, fig.cap = "Allometry models for LUMCON (6 years)."}

sub <- stemDat[(stemDat$region %in% "LUMCON") & (stemDat$status %in% "LIVE"), ]
fig2Col <- "gray40" # rgb(169, 169, 169, 140, maxColorValue = 255)
pointSize <- 0.3
tckSize <- 0.018
lwdSize <- 1.5
# png(filename = "allometry.png", width = 120, height = 120, units = "mm", res = 200)

par(mar = c(2, 2, 0.5, 0.5))
plot(sub$mass_tot ~ sub$hgt, 
     cex = pointSize, pch = 19, col = fig2Col,
     ylab = "", xlab = "", xlim = c(0, 172), ylim = c(0, 15.1), xaxt = "n", yaxt = "n",
     las = 1, tcl = 0.25, tck = tckSize, bty = "n", yaxs = "i", xaxs = "i")
abline(h = 0, lwd = lwdSize)
abline(v = 0, lwd = 1.5*lwdSize)
axis(side = 2, tcl = 0.25, tck = tckSize, at = axTicks(2), labels = FALSE, lwd = 1)
mtext(side = 2, text = "Mass (g)", line = 1.2)
mtext(side = 2, at = axTicks(2), text = axTicks(2), las = 1, line = 0.35)
axis(side = 1, tcl = 0.25, tck = tckSize, at = axTicks(1), labels = FALSE, lwd = 1)
mtext(text = axTicks(1), at = axTicks(1), side = 1, line = 0.2)
mtext(text = "Height (cm)", side = 1, line = 1.1)

x <- sub$hgt
y.pred2 <- models$coef.live[(models$region %in% "LUMCON") & (models$seas %in% "sprg")] * x^models$exp.live[(models$region %in% "LUMCON") & (models$seas %in% "sprg")]
y.pred3 <- models$coef.live[(models$region %in% "LUMCON") & (models$seas %in% "sumr")] * x^models$exp.live[(models$region %in% "LUMCON") & (models$seas %in% "sumr")]
y.pred4 <- models$coef.live[(models$region %in% "LUMCON") & (models$seas %in% "fall")] * x^models$exp.live[(models$region %in% "LUMCON") & (models$seas %in% "fall")]
y.pred5 <- models$coef.live[(models$region %in% "LUMCON") & (models$seas %in% "wint")] * x^models$exp.live[(models$region %in% "LUMCON") & (models$seas %in% "wint")] 
lines(x = x[order(y.pred2)], y = y.pred2[order(y.pred2)], lty = 1, lwd = 2)
lines(x = x[order(y.pred3)], y = y.pred3[order(y.pred3)], lty = 2, lwd = 2)
lines(x = x[order(y.pred4)], y = y.pred4[order(y.pred4)], lty = 3, lwd = 2)
lines(x = x[order(y.pred5)], y = y.pred5[order(y.pred5)], lty = 4, lwd = 2)
legend(x = 4, y = 13, legend = c("Spring", "Summer", "Fall", "Winter"), 
       lty = c(1:4), lwd = 2, bty = 'n', cex = 0.85)

# dev.off()

```


```{r allometry exponents and coefficients, fig.width = 3.5, fig.height = 2.5, warning=FALSE, echo=FALSE, fig.cap = "Relationship between allometry coefficients and exponents (one model for each marsh-season)."}

exponents_vs_coefs

```

The relationship between allometry exponents and coefficients remains intact and is consistent with Hill and Roberts (2017).


```{r stem demographics, fig.width = 6, fig.height = 4, echo=FALSE, warning=FALSE, fig.cap = "Temporal trends in stem size classes across all LUMCON marshes (mean +- SE of three marshes)."}

stm.demographics

```




## NAPP


```{r NAPP calculations, echo = FALSE, include = FALSE, message = FALSE}
### This section explores NAPP using the bioDat object
bioDat$moYr <- paste0(bioDat$mo, "-", substr(bioDat$yr, 3, 4))
bioDat$monthYear <- zoo::as.yearmon(bioDat$moYr, "%b-%y")
bioDat <- seasonLabel(data = bioDat)

marsh.biomass <- ddply(bioDat, .(region, site, yr, season, monthYear), summarise,
               above          =  mean(mass, na.rm = T),
               above.se       =  se(mass),
               stemDensity    =  mean(stems, na.rm = T),
               stemDensity.se =  se(stems),
               mass.dead      =  mean(mass.dead, na.rm = T), # stem biomass only
               mass.dead.se   =  se(mass.dead),              # stem biomass only
               litter         =  mean(litter, na.rm = TRUE),
               litter.se      =  se(litter),
               above.dead     =  mean(mass.dead, na.rm = T) + mean(litter, na.rm = TRUE), # stems + litter
               above.dead.se  =  sqrt(se(mass.dead)^2 + se(litter)^2),
               live.bg        =  mean(live.bg.gm2, na.rm = TRUE),
               live.bg.se     =  se(live.bg.gm2),
               dead.bg        =  mean(live.bg.gm2, na.rm = TRUE),
               dead.bg.se     =  se(dead.bg.gm2),
               tot.bg         =  mean(tot.bg.gm2, na.rm = TRUE),
               tot.bg.se      =  se(tot.bg.gm2)
)


### plot biomass over time
abv <- ggplot(marsh.biomass[marsh.biomass$region %in% "LUMCON", ], aes(y = above, x = monthYear)) + geom_pointrange(aes(ymax = above + above.se, ymin = above - above.se), size = 0.25) + theme_classic() + facet_grid(site ~ .) + labs(x = "", y = expression("Aboveground biomass (g "%.%m^-2*")"))


### NAPP
# "dead" column in napp is the sum of standing dead and litter
napp.full <- nappCalc(marsh.biomass, summarize = TRUE, EOS = TRUE, EOS_window = 0, annualReset = FALSE)
napp <- regionLabel(napp.full$summary)


napp.trend <- ggplot(napp[grepl(x = napp$site, pattern = "LUM"), ], aes(x = year, y = napp.smalley)) + 
  geom_bar(stat = "identity") + theme_classic() + facet_grid(. ~ site) + 
  labs(x = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), y = "")


### correlations between NAPP estimates
NAPP_columns <- c(5:6, 8:9, 17) # columns with NAPP estimates of interest




napp.melt <- melt(napp[grepl(x = napp$site, pattern = "LUM"), c(1, NAPP_columns)], id.vars = c("napp.smalley", "site"))
levels(napp.melt$variable) <- c("Milner and hughes 1968", "Valiela et al. 1975", "Peak biomass", "End-of-season")

napp.correlations <- ggplot(napp.melt, aes(x = napp.smalley, y = value, col = variable)) +
  geom_abline(slope = 1, intercept = 0, size = 0.7) + 
  labs(x = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), y = expression("NAPP (g "%.%m^-2%.%yr^-1~")")) + 
  theme_classic() + theme(legend.title = element_blank()) + 
  geom_point(aes(shape = as.factor(site))) + ylim(0, 3800) + xlim(0, 3400) + 
  geom_smooth(data = napp.melt, method = lm, se = TRUE, linetype = 1, size = 0.7, alpha = 0.2)

```


```{r NAPP drivers, echo = FALSE, include = FALSE, message = FALSE}
### This section explores environmental drivers of biomass and NAPP using the plot-level biomass data (bioDat object) and soil/water data (nutDat) 

### NAPP
dd.nut <- ddply(nutDat[nutDat$month.number %in% growingMonths, ], .(year, region, site), 
                numcolwise(mean, na.rm = TRUE))

napp.nuts <- join_all(list(napp, dd.nut))


### salinity relationships
rsqr.salinity <- round(summary(lm(napp.smalley ~ salinity.psu.bay, data = napp.nuts))$r.squared, 2)

rsqr.salinity.lum <- round(summary(lm(napp.smalley ~ salinity.psu.bay, data = napp.nuts[napp.nuts$region %in% "LUMCON", ]))$r.squared, 2)


napp.salt <- ggplot(napp.nuts, aes(x = salinity.psu.bay, y = napp.smalley)) + geom_point(aes(colour = site)) + theme_classic() + geom_smooth(method = "lm") + 
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creek salinity (psu)") # all data - looks nice

napp.salt.lum <- ggplot(napp.nuts[napp.nuts$region %in% "LUMCON", ], aes(x = salinity.psu.bay, y = napp.smalley, colour = site)) + geom_point() + theme_classic() + geom_smooth(method = "lm") + 
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creek salinity (psu)") # weak!

### NO3 relationship
rsqr.no3 <- round(summary(lm(napp.smalley ~ dissolved.no3.no2.bay, data = napp.nuts))$r.squared, 2)

rsqr.no3.lum <- round(summary(lm(napp.smalley ~ dissolved.no3.no2.bay, data = napp.nuts[napp.nuts$region %in% "LUMCON", ]))$r.squared, 2)


napp.no3 <- ggplot(napp.nuts, aes(x = I(log(dissolved.no3.no2.bay)), y = I(log(napp.smalley)))) + geom_point(aes(colour = site)) + theme_classic() + geom_smooth(method = "lm")  + 
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creekwater NO3 (uM)") # all data - looks pretty nice

napp.no3.lum <- ggplot(napp.nuts[napp.nuts$region %in% "LUMCON", ], aes(x = dissolved.no3.no2.bay, y = napp.smalley, colour = site)) + geom_point() + theme_classic() + geom_smooth(method = "lm") + 
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creekwater NO3 (uM)") # weak

```


```{r figure-aboveground biomass over time, fig.width = 4, fig.height = 4, echo=FALSE, message = FALSE, warning=FALSE, fig.cap = "Biomass at LUMCON marshes (mean +- SE of three plots)."}

abv

```


```{r figure-NAPP over time, fig.width = 6, fig.height = 3.5, echo=FALSE, warning=FALSE, fig.cap = "Smalley NAPP over time at three LUMCON marshes."}

napp.trend

```



```{r figure-correlations between NAPP estimates, fig.width = 5, fig.height = 4, echo=FALSE, warning=FALSE, fig.cap = "Correlations between NAPP estimates."}

napp.correlations
# ggsave("napp_correlations.png", width = 5.5, height = 4, units = "in")

```



```{r table-correlations between NAPP estimates, echo=FALSE, warning=FALSE, message = FALSE, fig.cap = "Correlations between NAPP estimates."}

# table with correlations between NAPP measures (for LUM only, all years)
# kable(corstarsl(as.matrix(napp[grepl(x = napp$site, pattern = "LUM"), NAPP_columns]))) # each site-year counted as independent obs

```




```{r figure-NAPP drivers, fig.width = 7, fig.height = 6, echo=FALSE, message = FALSE, warning=FALSE, fig.cap = "NAPP vs. salinity and creek NO3, including all data (top row) and LUMCON only (bottom row)."}

driver.fig <- arrangeGrob(napp.salt, napp.no3, napp.salt.lum, napp.no3.lum, nrow = 2)
plot(driver.fig)
```

The relationships between NAPP and bay salinity/NO3 are reasonably strong when Terrebonne Bay sites are included (r2 = 0.46 and 0.23 for salinity and NO3, respectively) but disappear when the dataset is reduced to LUMCON sites (r2 = 0.03 and 0.04 for salinity and NO3, respectively). 


## NDVI and Aboveground Biomass Relationship

```{r NDVI-biomass relationships, echo = FALSE, include = FALSE, message = FALSE}
#creating new dataset with 2017-2018 data only (since that is what NDVI data we have)
LUM<-subset(bioDat, region=="LUMCON")
LUM1718<-subset(LUM, yr > 2016)

#creating new dataset of aboveground biomass with means.
mean.mass<-summarySE(data=LUM1718, measurevar="mass", groupvars= c("site", "yr", "mo"))

#this is the NDVI data
# ndvi <- read_csv("ndvi_lum.csv")
# usethis::use_data(ndvi, overwrite = TRUE)


#Merge ndvi and lum data
mydata<-merge(ndvi, mean.mass, by=c("site","yr","mo"))

plot(avg.ndvi~mass, data=mydata)

rsqr.ndvi<- round(summary(lm(mass ~ avg.ndvi, data = mydata))$r.squared, 2)

mass.ndvi <- ggplot(mydata, aes(x = avg.ndvi, y = mass)) + geom_point(aes(colour = site)) + theme_classic() + 
  geom_smooth(method = "lm") + labs(x = ("NDVI"), y = expression("Aboveground biomass (g "%.%m^-2*")")) # all data - looks nice


##making text larger for figure
exp_vs_coefs <- ggplot(aes(y = exp.live, x = coef.live, colour = seas), data = models) + geom_point(size=3) + theme_classic() + labs(y = "exponent", x = "coefficient") +
  theme(text = element_text(size=14))
# spring = low coefficients, high exponents
##salt figure
napp.salt.lum <- ggplot(napp.nuts[napp.nuts$region %in% "LUMCON", ], aes(x = salinity.psu.bay, y = napp.smalley, colour = site)) + geom_point(size=3) + theme_classic() +  
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creek salinity (psu)") + theme(text = element_text(size=14))  # weak!


##no3 figure
napp.no3.lum <- ggplot(napp.nuts[napp.nuts$region %in% "LUMCON", ], aes(x = dissolved.no3.no2.bay, y = napp.smalley, colour = site)) + geom_point(size=3) + theme_classic() + 
  labs(y = expression("Smalley NAPP (g "%.%m^-2%.%yr^-1~")"), x = "Creekwater NO3 (uM)") + theme(text = element_text(size=14)) # weak


#stem density figure
live.stem<- summarySE(data=LUM, measurevar = "stems", groupvars=c("site", "yr", "mo"))
dead.stem<- summarySE(data=LUM, measurevar = "stems.dead", groupvars=c("site", "yr", "mo"))

```


```{r figure-NDVI_biomass, fig.width = 4, fig.height = 4, echo=FALSE, message = FALSE, warning=FALSE, fig.cap = "NDVI vs. aboveground biomass."}
mass.ndvi
#rsq is 0.33   

```

```{r figure-NDVI_biomass2, fig.width = 4, fig.height = 4, echo=FALSE, message = FALSE, warning=FALSE, fig.cap = "NDVI vs. aboveground biomass."}
# ggplot(live.stem,
#        aes(y = stems, x = as.numeric(time))) + #, col = region)) +
#   geom_point(size = 1.2) + geom_errorbar(aes(ymin = (stmDens - stmDens.se), ymax = (stmDens + stmDens.se)), width = 0) +
#   geom_line(lwd = 0.7) +
#   theme_classic() + theme(legend.title = element_blank()) +
#   ylab(expression("Live stem density (m"^-2~")")) + xlab("") +
#   facet_wrap("group", nrow = 2) +
#   theme_bw() %+replace% theme(strip.background  = element_blank(), legend.title = element_blank())
```


